Fast Neural Machine Translation Implementation

### 2018-05-18 wnmt.2018/wnmt.tex
Minor corrections, rephrasing, comments
Search for %!dtr! to find my comments.

  \author{Hieu Hoang$^{\dagger}$ \, Tomasz Dwojak$^{*}$ \, Rihards Krislauks$^{\ddagger}$ \\ {\bf Daniel Torregrosa$^{\mathparagraph}$ \, Kenneth Heafield$^{\dagger}$ } \\[2mm]
  $^{\dagger}$University of Edinburgh \, $^{*}$Adam Mickiewicz University in Pozna\'{n} \\ $^{\ddagger}$Tilde \,  $^{\mathparagraph}$Universitat d’Alacant }

@@

This paper describe the submissions to the efficiency track for GPUs by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante. We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation. We improve the performance with an efficient mini-batching algorithm and by fusing the softmax operation with the beam seach algorithm.

@@

We take an opposing approach by eschewing model selection in favour of efficient implementation. We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models. As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a means to explore the upper limit of the possible speed for a particular model and hardware. That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other considerations. %The consequence of this focused approach can be seen in the results of the shared task ???.

We submitted two systems to this year's shared task for the efficient translation on GPU. Our first submission was tailored to be as fast as possible while being above the baseline BLEU score. Our second submission trade some of the speed of first submission to return better quality translations.

@@

Most NMT more predict a large number of classes in their output layer, corresponding to the number of words or subword units in their target language. For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.

@@

Thirdly, the calculation of $max$ and $sum$ can be accomplished in one loop by adjusting $sum$ whenever a higher $max$ is found during the looping:
\begin{eqnarray*}
sum & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_b}} \\
    & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_a + \Delta}}  \\
    & = & e^{x_t - max_b} + e^{\Delta} \times \sum_{i=0...t-1}{e^{x_i - max_a}} 
\end{eqnarray*}
where $max_a$ is the original maximum value, $max_b$ is the new, higher maximum value, ie. $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.

@@

Both of our submitted systems uses the sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder. We use byte pair encoding~\citep{sennrich-haddow-birch:2016:P16-12} to adjust the vocabulary size.

@@

Our first system uses gated recurrent units (GRU) throughout, trained with Marian~\citep{junczys2018marian}, but submitted both systems using the Amun inference engine.

@@

Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, trained with an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models. We trained 2 systems with differing vocabulary sizes and varied the beam sizes before choosing the the settings that produces good translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.

@@

The efficiency of Amun's batching algorithm can be seen by observing the time taken for each decoding step in a batch of sentence, Figure~\ref{fig:decode-step}. Amun's decoding becomes faster as sentences finish translating. This contrast with the Marian inference engine, which uses a na\"ive batching algorithm, where the speed stays relatively constant throughout the decoding of the batch.

@@

Using batching can increase the translation speed by over 20 times in Amun, Figure~\ref{fig:batch-size}. Just as important, it doesn't suffer degradation with large batch sizes unlike the na\"ive algorithm which slows down when batch size over 1000 is used. This scalability issue is likely to increase in importance with new GPUs with ever increasing core counts.

@@

Fusing the bias and softmax operations in the output layer with the beam search results in a speed improvement by 25\%, Figure~\ref{fig:fused}. Its relative improvement decrease marginally with increasing beam size.

@@

Further insight can be gained by examining the time taken for each step in the output layer and beam search, Table~\ref{tab:fused-breakdown}. The fused operation only has to loop through the large cost matrix once, therefore, for low beam sizes its is comparable in speed to the simple kernel to add the bias. For higher beam sizes, the cost of maintaining the n-best list is felt.

@@

We have presented some of the improvement to Amun which are focused to improvement NMT inference.

>>>>

  \author{Hieu Hoang$^{\dagger}$ \, Tomasz Dwojak$^{*}$ \, Rihards Krislauks$^{\ddagger}$ \\ {\bf Daniel Torregrosa$^{\mathparagraph}$ \, Kenneth Heafield$^{\dagger}$ } \\[2mm]
  $^{\dagger}$University of Edinburgh \, $^{*}$Adam Mickiewicz University in Pozna\'{n} \\ $^{\ddagger}$Tilde \,  $^{\mathparagraph}$Universitat d'Alacant }

@@

This paper describes the submissions to the efficiency track for GPUs by members of the University of Edinburgh, Adam Mickiewicz University, Tilde and University of Alicante. We focus on efficient implementation of the recurrent deep-learning model as implemented in Amun, the fast inference engine for neural machine translation. We improve the performance with an efficient mini-batching algorithm and by fusing the softmax operation with the beam seach algorithm.

@@

We take an opposing approach by eschewing model selection in favour of efficient implementation. We use and enhanced a custom inference engine, Amun~\citep{junczys2016neural}, which we developed on the premise that fast deep-learning inference is an issue that deserves dedicated tools that are not compromised by competing objectives such as training or support for multiple models. As well as delivering on the practical goal of fast inference, it can serve as a test-bed for novel ideas on neural network inference, and is useful as a mean to explore the upper bound of the possible speed for a particular model and hardware. That is, Amun is an inference-only engine that supports a limited number of NMT models that put fast inference on modern GPU above all other considerations. %The consequence of this focused approach can be seen in the results of the shared task ???.

We submitted two systems to this year's shared task for the efficient translation on GPU. Our first submission was tailored to be as fast as possible while being above the baseline BLEU score. Our second submission trades some of the speed of first submission to return better quality translations.

@@

Most NMT models predict a large number of classes in their output layer, corresponding to the number of words or subword units in their target language. For example, ~\citet{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 sub-word units.

@@

Thirdly, the calculation of $max$ and $sum$ can be accomplished in one loop by adjusting $sum$ whenever a higher $max$ is found during the looping:
\begin{eqnarray*}
sum & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_b}} \\
    & = & e^{x_t - max_b} + \sum_{i=0...t-1}{e^{x_i - max_a + \Delta}}  \\
    & = & e^{x_t - max_b} + e^{\Delta} \times \sum_{i=0...t-1}{e^{x_i - max_a}} 
\end{eqnarray*}
where $max_a$ is the previous
%!dtr! original as in the one at the start of this iteration, or original as the one at the start of the process? 
maximum value, $max_b$ is the now higher maximum value, ie. $max_b > max_a$, and $\Delta = max_a - max_b$. The outline of our function is shown in Algorithm~\ref{algo:Fused Kernel}.

@@

Both of our submitted systems use the sequence-to-sequence model similar to that described in ~\citet{DBLP:journals/corr/BahdanauCB14}, containing a bidirectional RNN in the encoder and a two-layer RNN in the decoder. We use byte pair encoding~\citep{sennrich-haddow-birch:2016:P16-12} to adjust the vocabulary size.

@@

%Our first system uses gated recurrent units (GRU) throughout, trained with Marian~\citep{junczys2018marian}, but submitted both systems using the Amun inference engine.

Our first submitted system uses gated recurred units (GRU) throughout. It was trained using Marian~\citep{junczys2018marian}, but Amun was chosen as inference engine.

@@

Our second system uses multiplicative-LSTM~\citep{krause2017} in the encoder and the first layer of a decder, and a GRU in the second layer, trained with an extension of the Nematus~\citep{sennrich-EtAl:2017:EACLDemo} toolkit which supports such models. 
As with our first submission, Amun is used as inference engine.
We trained 2 systems with differing vocabulary sizes and varied the beam sizes, and chose the configuration that produced the best results for translation quality on the validation set, Table~\ref{tab:mLSTM BLEU - vocab sizes}.

@@

The efficiency of Amun's batching algorithm can be seen by observing the time taken for each decoding step in a batch of sentences, Figure~\ref{fig:decode-step}. Amun's decoding becomes faster as sentences are completely translated. This contrasts with the Marian inference engine, which uses a na\"ive batching algorithm, where the speed stays relatively constant throughout the decoding of the batch.

@@

Using batching can increase the translation speed by over 20 times in Amun, Figure~\ref{fig:batch-size}. Just as important, it doesn't suffer degradation with large batch sizes, unlike the na\"ive algorithm which slows down when batch sizes over 1000 is used. This scalability issue is likely to become more relevant as newer GPUs with ever increasing core counts are released.
%increase in importance with new GPUs with ever increasing core counts.

@@

Fusing the bias and softmax operations in the output layer with the beam search results in a speed improvement by 25\%, Figure~\ref{fig:fused}. Its relative improvement decreases marginally as the beam size increases.

@@

Further insight can be gained by examining the time taken for each step in the output layer and beam search, Table~\ref{tab:fused-breakdown}. The fused operation only has to loop through the large cost matrix once, therefore, for low beam sizes its is comparable in speed to the simple kernel to add the bias. For higher beam sizes, the cost of maintaining the n-best list is felt.
%!dtr! felt?

@@

We have presented some of the improvement to Amun which are focused on improving NMT inference.
%to improvement NMT inference.
